---
title: "Final Project"
author: "Vincent Smedile"
date: "13 December, 2020"
output: html_notebook
---

```{r,echo=FALSE,message=FALSE}
### Front Matter and Cleaning Environment
rm(list = ls())
```

## Introduction and Research Question

Throughout this year, there has been a consierable amount of discoursw surrounding what kind of states and economic systems can handle the pandemic best, especially given the comparatively low amount of deaths and infectios social democracies (such as Iceland and New Zealand) and marxist countries (such as Vietnam and Cuba) have experienced. As a result, it begs the question: Does state's socioeconomic and political structure (such as socialism, capitalism, the "Nordic Model, or Communism) effect it's ability to deal with the Coronavirus? And what other factors might cause the disparity we see between nations like the United States and nations like Vietnam? In order to do this, I will be categorizing nations into the following unique state types: Marxist (which would comprise of the four major states still espousing communism; China, Cuba, Laos, and Vietnam), Nordic Model States (which comprises what many academics consider to follow the unique "Nordic Third Way" of high economic freedoms and high social programs and collctive barganing such as Norway, Sweden, Iceland, Denmark, and Finland), Socialist Coalition Governments (which comprises of both dominant party republics and states with parliamentary elected socialist coalitions), and Capitalist countries (which comprises everyone else). To do this, I compared state type and a variable I created called "Covid Response Sucess Rate", or CRSR, which is a combined measure of a nations infection and death rates, and their total cases and deaths due to COVID. A lower score corresponds to a better COVID response, and is considered to be good. I will be separating the nations into different economic development tiers to ensure that there is no confounding influence of development. 

## Data

Below I have loaded the necessary packages for this analysis, including political science packages that will allow me to run preliminary regression analyses on my data.

```{r,message=FALSE}
# Load required packages
library(ggplot2)
library(rio)
library(descr)
library(stargazer)
library(dplyr)
```

### Sources of Data

The Primary source is the [COVID International Dataset](https://github.com/slinnpsu/PLSC309/blob/master/covid_international.csv), which was created by Clayton Webb of the Kansas State Unversity Political Science Department (a partner of my Political Science professor here at Penn State), who complied the dataset using datasets from the [Our World in Data](https://ourworldindata.org/) and [The Quality of Government Institute](https://qog.pol.gu.se/). I specifically obtained it through Dr. Suzanna Linn,  a Professor of Political Science here at Penn State (who is currently one of my other instructors). ISO Codes were taken from the [International Organization for Standardization](https://www.iso.org/about-us.html). The secondary source I obtained from the [World Bank](https://data.worldbank.org/indicator/NY.GNP.PCAP.CD). The Covid International Dataset is collected and maintained by Clayton Webb of KSU and the GNI data is collected and maintained by the World Bank. 
  
  The COVID International Dataset was collected by Professor Webb for the purposes of analyzing the social and political impacts of COVID-19 on international affairs. Due to his past partnerships with Professor Linn of the Penn State Political Science Department, the PSU Polisci Dept. also uses it for the same purposes and gives it to Polisci students for individual research practice. 

My main independent variable of analysis will be the **State_Type** variable, and my main dependent variable of analysis will be the **response_sucess_lev** variable, which corresponds to a nation's CRSR (COVID Response Sucess Rate)


```{r}
#Load Data Files

covid_data <- read.csv("covid_international.csv")
gni_data <- read.csv("Income ratings.csv")
```

### Overview of Data

Below I analyze the structure and composition, as well as cases of the COVID International Dataset and the GNI Data from the World Bank. Overallm it appears that a case constitutes a country in both datasets, andthat the location variable in the COVID data corresponds to a Country in the GNI data.

#### Analysis of the Covid International Dataset
```{r}
str(covid_data)
head(covid_data)
```


#### Analysis of the World Bank Data
```{r}
str(gni_data)
head(gni_data)
```

### Cleaning the Data and Creating the Variables of Analysis

Below, I will utilize data wrangling to separate and categorize each case (a country in the world) into one of the state type categories that I mentioned previously. I will also use this to construct the CRSR rating. 

```{r,message=FALSE}
# Here I will be correcting the death and infection rates for population, once I am done they will all be of the same unit (deaths per million people)
covid_data <-
  covid_data %>%
  mutate(death_rate = ((death_rate/population)*1000000))%>%
  mutate(inf_rate = ((inf_rate/population)*1000000))

# Below here I will be separating each nation into a specific state type by adding my own categorical variable "State Type" to each case. It is here that I will also code for a nations CRSR, which corresponds to the response_sucess_lev variable. 
marxist <-
  covid_data %>%
  filter(iso_code %in% c("CHN","LAO","CUB","VNM"))%>%
  select(location,iso_code,inf_rate,death_rate,deaths,cases,gdp_pc)%>%
  mutate(inf_prev_lev = inf_rate+cases)%>%
  mutate(death_prev_lev = deaths + death_rate)%>%
  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Marxist")

socialist_govs <-
  covid_data %>%
  filter(location %in% c("Nepal","Venezuela","Spain","New Zealand","Nicaragua","Argenti","Mozambique","Moldova","Angola","Algeria"))%>%
  select(location,iso_code,inf_rate,death_rate,deaths,cases,gdp_pc)%>%
  mutate(inf_prev_lev = inf_rate+cases)%>%
  mutate(death_prev_lev = deaths + death_rate)%>%
  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Socialist Coalition")

nordic_model <- 
  covid_data %>%
  filter(location %in% c("Iceland","Norway","Sweden","Denmark","Finland"))%>%
  select(location,iso_code,inf_rate,death_rate,deaths,cases,gdp_pc)%>%
  mutate(inf_prev_lev = inf_rate+cases)%>%
  mutate(death_prev_lev = deaths + death_rate)%>%
  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Nordic Model")

`%notin%` <- Negate(`%in%`) #allows me to not include the selected countries when filtering

capitalist_govs <- 
  covid_data %>%
  filter(location %notin% c("Iceland","Norway","Sweden","Denmark","Finland","Nepal","Venezuela","Spain","New Zealand","Nicaragua","Argenti","Mozambique","Moldova","Angola","Algeria"))%>%
  filter(iso_code %notin% c("CHN","LAO","CUB","VNM"))%>%
  select(location,iso_code,inf_rate,death_rate,deaths,cases,gdp_pc)%>%
  mutate(inf_prev_lev = inf_rate+cases)%>%
  mutate(death_prev_lev = deaths + death_rate)%>%
  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Capitalist")

# Below here I will be joining the individual datasets that contain each country type to create a complete dataset of the whole world's data. 

adapted_covid_data <-
  capitalist_govs %>%
    full_join(marxist)%>%
    full_join(nordic_model)%>%
    full_join(socialist_govs)%>%
    select(location,iso_code,state_type,response_sucess_lev,death_prev_lev,inf_prev_lev,gdp_pc)%>%
  arrange(location)

# Here I will finish the cleaning by joining the data from the COVID International Dataset with the World Bank's economic indicators. 
complete_cov_data <-
  adapted_covid_data %>%
  left_join(gni_data %>% select(Country.Code,IncomeGroup),by=c("iso_code"="Country.Code"))
```

<hr>

Here I will engage in an analysis of the structure of each dataset and clarify what each case means in this instance, as well as analyze the central tendancies, range, and other characteristics of the data. 

#### Analysis of the Now Cleaned Variables
```{r}
str(complete_cov_data)
head(complete_cov_data)
```

#### Central Tendancy, Range, Standard Dev., and Distribution: 

```{r}
complete_cov_data %>%
  ggplot(mapping=aes(x=response_sucess_lev))+
    geom_histogram(bins=20)+
    geom_vline(aes(xintercept=median(response_sucess_lev,na.rm=TRUE)),color="black",linetype="dashed")+
    labs(title="Figure 1: Distribution of Global Covid Response Ratings",
         caption="Source: COVID International Dataset",
         x="COVID Response Sucess Rating",
         y="Frequency")
```

```{r}
complete_cov_data%>%
  ggplot(mapping=aes(x=response_sucess_lev,facet=factor(state_type),fill=factor(state_type)))+
  geom_density(alpha=.5,aes(color=factor(state_type)))+
  guides(color=FALSE)+
  labs(title="Density Distribution of COVID Response Rating \nby State Type",
       caption="Source: COVID International Dataset \nWorld Bank",
       fill="State Type",
       x="Covid Response Sucess Rating",
       y="Density")+
  scale_x_continuous(limits=c(0,15000))+
  facet_wrap(~state_type)
```

```{r}
complete_cov_data%>%   
  ggplot(mapping=aes(x=state_type,y=response_sucess_lev,fill=state_type))+
  geom_boxplot()+
  labs(title="Distribution of COVID Response Ratings by State Type",
       caption="Source: COVID International Dataset \nWorld Bank",
       x=NULL,
       y="COVID Response Rating")+
  scale_y_continuous(limits=c(0,10000),breaks=c(0,1000,2000,3000,4000,5000,6000,7000,8000,9000,10000))+
  scale_fill_discrete(name="State Type")
```

#### For capitalist States
```{r}
median(capitalist_govs$response_sucess_lev,na.rm=TRUE)
sd(capitalist_govs$response_sucess_lev,na.rm=TRUE)
range(capitalist_govs$response_sucess_lev,na.rm=TRUE)
nrow(capitalist_govs)
```

<hr>

#### For Socialist Coalition Govs
```{r}

median(socialist_govs$response_sucess_lev,na.rm=TRUE)
sd(socialist_govs$response_sucess_lev,na.rm=TRUE)
range(socialist_govs$response_sucess_lev,na.rm=TRUE)
nrow(socialist_govs)
```

<hr>

#### For Nordic nations
```{r}

median(nordic_model$response_sucess_lev,na.rm=TRUE)
sd(nordic_model$response_sucess_lev,na.rm=TRUE)
range(nordic_model$response_sucess_lev,na.rm=TRUE)
nrow(nordic_model)
```

<hr>

#### For Marxist States
```{r}

median(marxist$response_sucess_lev,na.rm=TRUE)
sd(marxist$response_sucess_lev,na.rm=TRUE)
range(marxist$response_sucess_lev,na.rm=TRUE)
nrow(marxist)
```

<hr>

#### For world data
```{r}

median(complete_cov_data$response_sucess_lev,na.rm = TRUE)
sd(complete_cov_data$response_sucess_lev,na.rm = TRUE)
range(complete_cov_data$response_sucess_lev,na.rm = TRUE)
nrow(complete_cov_data)
```

<hr>

From the data analysis, the central tendency of around 8108.77 (computed using the median of the world’s CRSRs), with a standard deviation of around 533989 and a range of 1.927102e+01 to 5.573396e+06. The distribution of CRSRs exhibited a unimodal distribution of CRSRs skewed to the right, signifying that the nations generally had low CRSRs compared to one another, with a sizable degree of dispersion due to the presence of glaring outliers (possibly the result of nations that chose to ignore or were unable to combat the pandemic due to a lack of resources). It is also worth noting that the Nordic Model nations appeared to have the lowest central tendency, with a median CRSR of 36575.45. Marxist states were the next lowest with a central tendency of around 46325.37, and socialist coalition governments the next lowest with 130061.4. Capitalist governments had a much higher median of 564849.8. The ranges of CRSR values varies for each state type varies, however. The Nordic Model countries appear to have the highest minimum and maximum values, and the highest mean CRSR. Marxist states had the lowest range of values, as well as the lowest mean CRSR rating. Capitalist governments had the next lowest mean and had a lower minimum and maximum than the socialist coalition and Nordic state types. Lastly, socialist coalition governments had a minimum, maximum, and mean smaller only than the Nordic Model states. 

<hr>

### Analysis of the Relationship Between Variables and Results


#### Concluding Remarks 

