---
title: "R Notebook"
author: "Vincent Smedile"
date: "15 November, 2020"
output: html_notebook
---

### Front Matter
```{r}
# Load required packages
library(ggplot2)
library(rio)
install.packages("descr")
library(descr)
install
library(sjPlot)
library(stargazer)

#Load Data Files

covid <- import("covid_international.csv")
income <- import("Income ratings.csv")
```


### Preliminary Cleaning of Data

```{r}

```


###Guiding Question (i.e., Research Question): Make a clear statement of a single guiding/research question that will drive your analysis.  It should be phrased as a question that you will try to investigate (i.e., it should have a question mark) 

#### - Become acquainted with your data sources

  - Where did you find them?
  
  
  
  - Who collected/maintains them?
  
  
  
  - When & Why were they originally collected?
  
  
  
  - What does a case represent in each data source, and how many total cases are available?
  
  
  
  - What are some of the variables that you plan to use?


### Explore intuition related to the research question 

#### 1. Create some informative plots and summary statistics 


#### 2. Describe preliminary observations and intuition about the research question

