---
title: "Final Project"
author: "Vincent Smedile"
date: "13 December, 2020"
output: html_notebook
---

```{r,echo=FALSE,message=FALSE}
### Front Matter and Cleaning Environment
rm(list = ls())
```

## Introduction and Research Question

Throughout this year, there has been a consierable amount of discoursw surrounding what kind of states and economic systems can handle the pandemic best, especially given the comparatively low amount of deaths and infectios social democracies (such as Iceland and New Zealand) and marxist countries (such as Vietnam and Cuba) have experienced. As a result, it begs the question: Does state's socioeconomic and political structure (such as socialism, capitalism, the "Nordic Model, or Communism) effect it's ability to deal with the Coronavirus? And what other factors might cause the disparity we see between nations like the United States and nations like Vietnam? In order to do this, I will be categorizing nations into the following unique state types: Marxist (which would comprise of the four major states still espousing communism; China, Cuba, Laos, and Vietnam), Nordic Model States (which comprises what many academics consider to follow the unique "Nordic Third Way" of high economic freedoms and high social programs and collctive barganing such as Norway, Sweden, Iceland, Denmark, and Finland), Socialist Coalition Governments (which comprises of both dominant party republics and states with parliamentary elected socialist coalitions), and Capitalist countries (which comprises everyone else). To do this, I compared state type and a variable I created called "Covid Response Sucess Rate", or CRSR, which is a combined measure of a nations infection and death rates, and their total cases and deaths due to COVID. A lower score corresponds to a better COVID response, and is considered to be good. I will be separating the nations into different economic development tiers to ensure that there is no confounding influence of development. 

## Data

Below I have loaded the necessary packages for this analysis, including political science packages that will allow me to run preliminary regression analyses on my data.

```{r,message=FALSE}
# Load required packages
library(ggplot2)
library(rio)
library(descr)
library(stargazer)
library(dplyr)
```

### Sources of Data

The Primary source is the [COVID International Dataset](https://github.com/slinnpsu/PLSC309/blob/master/covid_international.csv), which was created by Clayton Webb of the Kansas State Unversity Political Science Department (a partner of my Political Science professor here at Penn State), who complied the dataset using datasets from the [Our World in Data](https://ourworldindata.org/) and [The Quality of Government Institute](https://qog.pol.gu.se/). I specifically obtained it through Dr. Suzanna Linn,  a Professor of Political Science here at Penn State (who is currently one of my other instructors). ISO Codes were taken from the [International Organization for Standardization](https://www.iso.org/about-us.html). The secondary source I obtained from the [World Bank](https://data.worldbank.org/indicator/NY.GNP.PCAP.CD). The Covid International Dataset is collected and maintained by Clayton Webb of KSU and the GNI data is collected and maintained by the World Bank. 
  
  The COVID International Dataset was collected by Professor Webb for the purposes of analyzing the social and political impacts of COVID-19 on international affairs. Due to his past partnerships with Professor Linn of the Penn State Political Science Department, the PSU Polisci Dept. also uses it for the same purposes and gives it to Polisci students for individual research practice. 

My main independent variable of analysis will be the **State_Type** variable, and my main dependent variable of analysis will be the **response_sucess_lev** variable, which corresponds to a nation's CRSR (COVID Response Sucess Rate)


```{r}
#Load Data Files

covid_data <- read.csv("covid_international.csv")
gni_data <- read.csv("Income ratings.csv")
```

### Overview of Data

Below I analyze the structure and composition, as well as cases of the COVID International Dataset and the GNI Data from the World Bank. Overallm it appears that a case constitutes a country in both datasets, andthat the location variable in the COVID data corresponds to a Country in the GNI data.

#### Analysis of the Covid International Dataset
```{r}
str(covid_data)
head(covid_data)
```


#### Analysis of the World Bank Data
```{r}
str(gni_data)
head(gni_data)
```

### Cleaning the Data and Creating the Variables of Analysis

Below, I will utilize data wrangling to separate and categorize each case (a country in the world) into one of the state type categories that I mentioned previously. I will also use this to construct the CRSR rating. 

```{r,message=FALSE}
# Here I will be correcting the death and infection rates for population, once I am done they will all be of the same unit (deaths per million people)
covid_data <-
  covid_data %>%
  mutate(inf_rate_mod = (inf_rate/wdi_pop)*1000000)%>%
  mutate(death_rate_mod = (death_rate/wdi_pop)*1000000)

# Below here I will be separating each nation into a specific state type by adding my own categorical variable "State Type" to each case. It is here that I will also code for a nations CRSR, which corresponds to the response_sucess_lev variable. 
marxist <-
  covid_data %>%
  filter(iso_code %in% c("CHN","LAO","CUB","VNM"))%>%
  select(location,iso_code,inf_rate_mod,death_rate_mod,deaths,cases,gdp_pc,pop_density,response, fh_polity2)%>%
  mutate(inf_prev_lev = inf_rate_mod+cases)%>%
  mutate(death_prev_lev = deaths + death_rate_mod)%>%
  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Marxist")

socialist_govs <-
  covid_data %>%
  filter(location %in% c("Nepal","Venezuela","Spain","New Zealand","Nicaragua","Argenti","Mozambique","Moldova","Angola","Algeria","Zimbabwe"))%>%
  select(location,iso_code,inf_rate_mod,death_rate_mod,deaths,cases,gdp_pc,pop_density,response, fh_polity2)%>%
  mutate(inf_prev_lev = inf_rate_mod+cases)%>%
  mutate(death_prev_lev = deaths + death_rate_mod)%>%

  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Socialist Coalition")

nordic_model <- 
  covid_data %>%
  filter(location %in% c("Iceland","Norway","Sweden","Denmark","Finland"))%>%
  select(location,iso_code,inf_rate_mod,death_rate_mod,deaths,cases,gdp_pc,pop_density,response, fh_polity2)%>%
  mutate(inf_prev_lev = inf_rate_mod+cases)%>%
  mutate(death_prev_lev = deaths + death_rate_mod)%>%
  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Nordic Model")

`%notin%` <- Negate(`%in%`) #allows me to not include the selected countries when filtering

capitalist_govs <- 
  covid_data %>%
  filter(location %notin% c("Iceland","Norway","Sweden","Denmark","Finland","Nepal","Venezuela","Spain","New Zealand","Nicaragua","Argenti","Mozambique","Moldova","Angola","Algeria","Zimbabwe"))%>%
  filter(iso_code %notin% c("CHN","LAO","CUB","VNM"))%>%
  select(location,iso_code,inf_rate_mod,death_rate_mod,deaths,cases,gdp_pc,pop_density,response, fh_polity2)%>%
  mutate(inf_prev_lev = inf_rate_mod+cases)%>%
  mutate(death_prev_lev = deaths + death_rate_mod)%>%
  mutate(response_sucess_lev = inf_prev_lev+death_prev_lev)%>%
  mutate(state_type = "Capitalist")

# Below here I will be joining the individual datasets that contain each country type to create a complete dataset of the whole world's data. 

adapted_covid_data <-
  capitalist_govs %>%
    full_join(marxist)%>%
    full_join(nordic_model)%>%
    full_join(socialist_govs)%>%
    select(location,iso_code,state_type,response_sucess_lev,death_prev_lev,inf_prev_lev,gdp_pc,pop_density,response,fh_polity2)%>%
  arrange(location)

# Here I will finish the cleaning by joining the data from the COVID International Dataset with the World Bank's economic indicators. 
complete_cov_data <-
  adapted_covid_data %>%
  left_join(gni_data %>% select(Country.Code,IncomeGroup),by=c("iso_code"="Country.Code"))
complete_cov_data
```

<hr>

Here I will engage in an analysis of the structure of each dataset and clarify what each case means in this instance, as well as analyze the central tendancies, range, and other characteristics of the data. 

#### Analysis of the Now Cleaned Variables
```{r}
str(complete_cov_data)
head(complete_cov_data)
```

<hr>

#### Central Tendancy, Range, Standard Dev, and Distribution of CRSRs: 

The dependent variable will be defined as “COVID Response Success,” which will be comprised of two components: Infection Prevention Level and Death Prevention Level. Infection Prevention Level will be an interval variable comprised of a summation of the Average COVID Cases per day (the average daily infection rate) and the Total Number of COVID Cases (total number of COVID deaths during the pandemic) through August 17th 2020. Death Prevention Level will be an interval variable as well, which is created by a summation of the Average COVID Deaths per Day (the average deaths in a country over the course of the pandemic) and the Total number of COVID Deaths (deaths due to COVID over course of pandemic) .To determine each nation’s COVID Response Success Rating (CRSR), Infection Prevention and Death Prevention will be added together to form one singular score. Nations with the lowest score will be considered to have a better success at dealing with the COVID-19 pandemic.

From the data analysis, the central tendency or median of around 8108.77 (computed using the median of the world’s CRSRs), with a standard deviation of around 533989 and a range of 19 to 5.57 million. The distribution of CRSRs exhibited a unimodal distribution of CRSRs skewed to the right, signifying that the nations generally had low CRSRs compared to one another, with a sizable degree of dispersion due to the presence of glaring outliers (possibly the result of nations that chose to ignore or were unable to combat the pandemic due to a lack of resources).

<hr>

```{r}
complete_cov_data %>%
  ggplot(mapping=aes(x=response_sucess_lev))+
    geom_histogram(bins=20)+
    geom_vline(aes(xintercept=median(response_sucess_lev,na.rm=TRUE)),color="black",linetype="dashed")+
    labs(title="Figure 1: Distribution of Global Covid Response Ratings",
         caption="Source: COVID International Dataset",
         x="COVID Response Sucess Rating",
         y="Frequency")
```

```{r}
complete_cov_data%>%
  ggplot(mapping=aes(x=response_sucess_lev,facet=factor(state_type),fill=factor(state_type)))+
  geom_density(alpha=.5,aes(color=factor(state_type)))+
  guides(fill=FALSE,color=FALSE)+
  labs(title="Density Distribution of COVID Response Rating \nby State Type",
       caption="Source: COVID International Dataset \nWorld Bank",
       fill="State Type",
       x="Covid Response Sucess Rating",
       y="Density")+
  scale_x_continuous(limits=c(0,15000))+
  facet_wrap(~state_type)
```

```{r}
complete_cov_data%>%   
  ggplot(mapping=aes(x=state_type,y=response_sucess_lev,fill=state_type))+
  geom_boxplot()+
  guides(fill=FALSE,color=FALSE)+
  labs(title="Distribution of COVID Response Ratings by State Type",
       caption="Source: COVID International Dataset \nWorld Bank",
       x=NULL,
       y="COVID Response Rating")+
  scale_y_continuous(limits=c(0,10000),breaks=c(0,1000,2000,3000,4000,5000,6000,7000,8000,9000,10000))+
  scale_fill_discrete(name="State Type")
```

```{r}
complete_cov_data %>%
  filter(!is.na(IncomeGroup)) %>%
  ggplot(mapping=aes(x=state_type,y=response_sucess_lev,fill=state_type))+
  geom_boxplot()+
  guides(fill=FALSE,color=FALSE)+
  labs(title="Distribution of COVID Response Ratings by State Type",
       caption="Source: COVID International Dataset \nWorld Bank",
       x=NULL,
       y="COVID Response Rating")+
  facet_wrap(~IncomeGroup)+
  scale_y_continuous(limits=c(0,10000),breaks=c(0,1000,2000,3000,4000,5000,6000,7000,8000,9000,10000))+
  scale_fill_discrete(name="State Type")
```


#### For capitalist States
```{r}
median(capitalist_govs$response_sucess_lev,na.rm=TRUE)
sd(capitalist_govs$response_sucess_lev,na.rm=TRUE)
range(capitalist_govs$response_sucess_lev,na.rm=TRUE)
nrow(capitalist_govs)
```

<hr>

#### For Socialist Coalition Govs
```{r}

median(socialist_govs$response_sucess_lev,na.rm=TRUE)
sd(socialist_govs$response_sucess_lev,na.rm=TRUE)
range(socialist_govs$response_sucess_lev,na.rm=TRUE)
nrow(socialist_govs)
```

<hr>

#### For Nordic nations
```{r}

median(nordic_model$response_sucess_lev,na.rm=TRUE)
sd(nordic_model$response_sucess_lev,na.rm=TRUE)
range(nordic_model$response_sucess_lev,na.rm=TRUE)
nrow(nordic_model)
```

<hr>

#### For Marxist States
```{r}

median(marxist$response_sucess_lev,na.rm=TRUE)
sd(marxist$response_sucess_lev,na.rm=TRUE)
range(marxist$response_sucess_lev,na.rm=TRUE)
nrow(marxist)
```

<hr>

#### For world data
```{r}

median(complete_cov_data$response_sucess_lev,na.rm = TRUE)
sd(complete_cov_data$response_sucess_lev,na.rm = TRUE)
range(complete_cov_data$response_sucess_lev,na.rm = TRUE)
nrow(complete_cov_data)
```

<hr>

#### Central Tendancy, Range, Standard Dev, and Distribution of State Types:

The independent variable will be defined as “State Type,” which will be a selection of cases handpicked and categorized by myself from the variable “country” from the COVID International Data Set. State Type will be a categorical variable broken down COVID the following five types. Nordic Social Democracies, which include nations that follow the “Nordic Model” (Iceland, Finland, Sweden, Denmark, and Norway) (Simon, 2017). Socialist Coalition Governments, or nations currently governed by an elected leftist party (such as New Zealand, Venezuela, Nepal, or Spain). State Socialist Nations, which are four of the five remaining nations espousing socialism, communism, Marxism (People’s Republic of China, Vietnam, Cuba, and Laos). Lastly, all other nations that are not specifically organized into the following categories will be considered “free-market capitalist” states.

The preliminary analysis of the State Type variable displayed that most of the world is comprised of nations that fit into the “Free Market Capitalist” category, with what can be considered a central tendency toward capitalism and a unimodal distribution when it comes to socioeconomic type for states in the 21st century. Overall, there were 194 capitalist countries, 11 socialist coalition governments, 5 Nordic Model countries, and only 4 Marxist states. The large array of nations from different income rankings, continents, conditions, etc. within the capitalist category would account for the great deal of variation and outliers present within the capitalist states, which was lacking in the other state type variables.

<hr>

```{r}
complete_cov_data %>%
  ggplot(mapping=aes(x=state_type,fill=state_type))+
    geom_bar()+
    guides(fill=FALSE,color=FALSE)+
    labs(title="Distribution of State Types",
         x="State Type",
         y="Total Countries")+
  scale_fill_discrete(name="State Type")
```

### Analysis of the Relationship Between Variables and Results

<hr>
#### Preliminary Analysis of the Variables

As shown in Figure 3, when separated by State Type, this analysis' main independent variable, it was seen that generally, the Marxist states performed the best, having both the lowest minimum and maximum CRSRs of around 20 to 3500, and the lowest median CRSR of around 1000. Interestingly, the capitalist states (despite having the largest range and maximum value), had the second lowest median CRSR of around 1600. Socialist Coalition governments had the third lowest median of around 2800, followed by Nordic Model States with a median of 5000. However, as seen in Figure 4 due to the overlap of the error bars, at the moment this relation indicates that there is not a statistical relationship between State Type and CRSRs, and a p-value > .05.

```{r}
complete_cov_data%>%
  filter(!is.na(state_type) & !is.na(response_sucess_lev))%>%
  group_by(state_type)%>% 
  summarize(mean=mean(response_sucess_lev, na.rm=TRUE), n=n(), sd=sd(response_sucess_lev, na.rm=TRUE))%>%
  mutate(margin=1.96*sd/sqrt(n)) %>%
ggplot(mapping=aes(y=mean, x=factor(state_type))) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=mean-margin, ymax=mean+margin), width=.1)+
  labs(title="Figure 4: Correlation Analysis Between State Type and CRSR",
       x=NULL,
       y= "COVID Response Rating")
```

<hr>

#### Multiple Regression Analysis of the Variables

A linear regression model was conducted to determine the relationship between the variables. Overall, there was no statistically significant relationship found between State Type and CRSRs, although all leftist states types did create a decrease of CRSR.

The following Variables were utilized as controls:

- **Economic Development Level:** Countries with a higher income level naturally would have access to better resources and be better equipped to fight the pandemic.
- **Population Density:** Countries with more persons closely packed could have a higher viral spread.
- **Democracy Rating:** A lower democracy index rating might correlate to the citizens of the nation having restricted lives anyways, which could lower COVID cases.
- **Stringency of Response:** Nations with tighter lockdowns and more closures of public spaces may have lower transmission of COVID outside their socialized medicine and economic apparatus.

```{r}
multi_reg1 <- lm(response_sucess_lev ~ state_type + IncomeGroup+ pop_density + response + fh_polity2, data=complete_cov_data)
summary(multi_reg1)
```


Overall, each variable had the following effects on a country's CRSR when all other variables are held constant. Note that Capitalist states and Upper Income nations are not included as they were used as the reference/constant for the Income and State Type factor:

- **Marxist:** CRSR value will drop around 59742 points, which means being a Marxist state translates to handling COVID better.
- **Socialist Coalition Countries:** the CRSR drops on average around 250746 points, meaning having a socialist coalition or party in power enacting such policies translates to handling COVID better.
- **Nordic Model Countries:** the CRSR value drops 83740 points, meaning the Nordic Model practiced by the Nordic countries increases a government's ability to handle COVID.
- **Lower Income Nations:** It also appears that Low Income are likely to have lower CRSRs with a predicted drop in around 104565
- **Lower Middle Income Nations:** Appear to decrease CRSRs 31894 points.) for both.
- **Upper Middle Income:** Being an upper middle income nation causes a predicted 94818 point increase in CRSR, implying that wealthier nations actually do worse.
- **Response Stringency:** States with a higher response stringency were likely to have a lower CRSR, with every 1 point higher on the Response Stringency Index amounting to a predicted 1848.94 point drop in CRSR.
- **Democracy Index Rating:** The higher one scores on a scale of 1-10, the higher the CRSR. For every 1 point increase on the combined Freedom House-Polity Democracy score (meaning more democratic), there is a predicted 20650.32 point increase in CRSR. Meaning that the more democratic a nation is, the worse it's ability to handle the COVID pandemic.


```{r}
library(stargazer)
stargazer(multi_reg1,
          type='text',
          title = "Multi Regression of CRSR Sucess Ratings",
          dep.var.caption = "COVID Response Sucess Rating",
          covariate.labels = c("Marxist State","Nordic Model","Socialist Coalition","Low Income","Lower Middle Income","Upper Middle Income","Population Density","Response Stringency","Democracy Rating","Intercept"))
```

#### Concluding Remarks 

Based on the data analysis presented herein, it can be said with some degree of certainty that the socioeconomic organization of a state does not affect it’s ability to better combat the pandemic. As shown by the predictive modeling, the average CRSR for Socialist Coalition States, Marxist States, and Nordic Model States were all lower than the CRSR for capitalist states, which translates to a better response as the CRSR is a combined measure that denotes smaller deaths, infection count, mortality rate due to COVID, and smaller infection rate. Alongside this was possible predictive evidence to suggest that nations with less traditional liberal democratic structures might be better equipped at dealing with the pandemic. This could possibly be due in part to the social ostracization and collectivist mindset usually incorporated into authoritarian regimes. 

Furthermore, there is the possibiity that the study was constrained by sources of error and limitations. The model itself does not have a very high level of confidence (low $R^2$ value), thus leading to the possibility that the relation we are seeing is indeed real, but requires further analysis of possible confounding and control variables in order to determine if this prediction is in fact real. Moreover, due to the possible evidence of the suppression of counting and data collection in places such as the United States or China, there is a chance that the numbers presented are not correct and thus prevent a proper relationship. It is possible that there are cultural aspects and social aspects, such as social traditions that may not be accounted for or could better predict how leftist policies affect COVID., for example, the Confucianist traditions of China and Vietnam and the cultural collectivism inherent in Nordic and Viking culture might be a confounding variable. Another possible issue is that the categorization for State Type, specifically the Socialist Coalition category, was assigned based on personal knowledge and research. Furthermore, parties with officially leftwing ideas may not always enact the socialistic policies shown in past literature to better deal with pandemics (see Sweden and Spain's comparatively center to center-right wing responses to COVID). 